Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-24 supplement.tex
Fix app refs (partially)

\section{Immediate benefits of positive design for negative design}

Already previous work on RNA Design, e.g. INFO-RNA and Incarnation, observed
that---put briefly---negative design benefits from positive design. Here, more
concretely, we simply \emph{demonstrate} this effect (for small examples), where
we consider \Def{negative design} as  
designing RNA sequences that fold with high probability $Pr(R|S)$ into the
given target structure $R$; and \Def{positive design}, as generating one or
many sequences that have a low energy for $R$.

@@

Note that low energy $E(S|R)$ is (non-linearly) correlated with high
probability $Pr(S|R)$, the probability to see $S$ in the Boltzmann distribution
of sequences folding into $R$. Recall that the probability $Pr(R|S)$ can be
calculated as $\exp(-E(S,R)/RT)/Z(S)$,
where $Z(S)=\sum_{R'} \exp(-E(S,R')/RT)$.
%
Consequently, Figure~\ref{fig:negforpos} ultimately illustrates the dependencies
between $Pr(S|R)$ and $Pr(R|S)$ for uniform samples and low energy samples
(where those samples can be efficiently generated due to our Boltzmann sampling
strategy). For our examples, we observe (Fig.~\ref{fig:negforpos}) that Boltzmann sampling with weight 500 (green), generates good negative designs 
with substantially higher probability than uniform sampling (blue).

@@

\begin{figure}[h!]
\textbf{A}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/stem_weight1-500_mean}%
\textbf{B}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight1-500_mean}%
\caption{Comparison of uniform sampling and Boltzmann sampling with
weight $\pi=500$. We show results for two different small structures: (A) single stem structure (B) double stem structure. For each respective structure $R$, we plot the energies $E(S,R)$ of
10000 sampled sequences against the probabilities $Pr(R|S)$ (in log scale). For illustrating the trend, we show moving averages over windows of $\pm 1$ kcal. Note that while only the energy $E(S,R)$ is directly controlled by Boltzmann sampling, this strongly impacts the common negative design objective $Pr(R|S)$.}
\label{fig:negforpos}
\end{figure}

@@

\begin{figure}[h!]
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering\includegraphics[width=0.8\textwidth]{Figs/td-widths}
  \caption{Treewidths for multi-target RNA design instances of
    different complexity. Distributions of treewidths shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-widths}
\end{figure}

\begin{figure}[h!]
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering
  \includegraphics[width=0.8\textwidth]{Figs/td-times}
  \caption{Computation time of tree decompositions for
    multi-target RNA design instances of different complexity.
    Distributions of times (in ms/instance) shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-times}
\end{figure}

For each instance, we generated the dependency graphs in the base pair
model and in the stacking model. Then, we performed tree decomposition
(using strategy ``GreedyFillIn'' of LibTW~\citep{Dijk2006}) on each dependency
graph. The obtained treewidths are reported in
Fig.~\ref{fig:td-widths}, while Fig.~\ref{fig:td-times} shows the
corresponding run-times of the tree decomposition.

\begin{figure}[h!]
  %% the figure shows benchmark_inputs/RNAfold/3str/f3.100.0.inp
  \centering
  {\small \verbatiminput{Figs/f3_100_0.txt}}
  \includegraphicstop[width=0.55\textwidth]{Figs/td-example-basepair}%
  \includegraphicstop[width=0.30\textwidth]{Figs/td-example-stacking}
  \caption{Example instance from RF3 (top) with its tree decompositions in the base pair (left) and stacking model (right). The respective treewidths are 2 and 12.}
  \label{fig:td-example}
\end{figure}

\revised{
  Finally, for further illustration, we show concrete tree decompositions for a selected instance in the benchmark set RF3 in Figure~\ref{fig:td-example}. This figure is provided solely for illustrational purposes:
  First, it depicts the tree decomposition for a typical, realistic sampling task. Second, it exemplarily illustrates the typical impact of the energy model on the tree decomposition.
  Compared to the tree decomposition based on the stacking model, the tree decomposition for the base pair model is much simpler. This shows in many connected components, which correspond to independently solvable sub-problems, and fewer positions per bag (i.e. smaller tree width), which allows far more efficient evaluation per tree node. 
  Note that the complexity of the tree decomposition ultimately shows the complexity of the dependencies that need to be considered for simultaneous energy evaluation for all target structures. This complexity is substantially increased in the stacking model over the simpler base pair model.
}

@@

\section{Run times for generating samples}

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times (Intel(R) Core(TM) i7-4770 CPU) for generating 10000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant pre-computation, it is informative to separately show construction and sampling times (Figure~\ref{fig:run-times-redprint}). Moreover, \ourprog{} can be used with a simpler base pair model or a rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width and thus the runtime, as noticeable in Figure~\ref{fig:run-times-redprint}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \ourprog{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences. }
  \label{fig:run-times-redprint}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \RNAblueprint{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences.}
  \label{fig:run-times-rbp}
\end{figure}

A fair comparison to sampling times by other tools can hardly be given, since the other tools produce samplings of very different nature. In particular, Modena does not make any guarantees on properties of the sampling. Non-surprisingly, this allows it to sample significantly faster (several orders of magnitudes).
Nevertheless, the comparison to run times of \RNAblueprint{} is still interesting, since this tool generates---similarly to our tool---\emph{controlled} samples. As we argue that the Boltzmann sampling is fundamentally more flexible than uniform sampling as generated by \RNAblueprint, \ourprog{} performs even particularly well, beating \RNAblueprint{} most of the time with only a few exceptions for particularly hard instances. The results for \RNAblueprint{} are shown in Figure~\ref{fig:run-times-rbp}.

@@

\section{Illustrating multi-dimensional Boltzmann sampling for three pseudoknot-free target structures}

    In analogy to Fig.~\ref{fig:energydist-pk} of the main text,
    Fig.~\ref{fig:energydist} illustrates our multi-dimensional sampling
    strategy for three pseudoknot-free target structures (example instance
    from the \texttt{Modena} benchmark). The figure shows Turner energy
    distributions of single structures from sampling with different weight
    parameters, where weights are trained to target specific energies. For
    comparison, the figure shows distributions from uniform and Boltzmann
    sampling at respective homogenous weights $1$ and $e^\beta$.

    Fig.~\ref{fig:gc-content} shows the distributions of the GC content of the samples, while targeting different combinations of energies in this experiment. For completeness, we also report corresponding results for the analogous experiment from the main text. We observe that in addition to the target energies, also the feature $\GCb\%$ is narrowly defined by the multi-dimensional Boltzmann sampling strategy. In comparison, for uniform sampling and Boltzmann sampling at high weight $e^\beta$, where the \GCb\% is not controlled, we obtain much broader distributions.

\begin{figure}[h!]
      \begin{center}
        %{\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
        %{\sf \bfseries B}
        \includegraphicstop[width=0.8\textwidth]{Figs/energy_distribution}
      \end{center}
      \caption{%
        Targeting specific energies using multi-dimensional Boltzmann
        sampling. We show the Turner energy distributions for
        three target structures R1, R2 and R3. By
        targeting the respective free energies $(-40,-40,-20)$ and
        $(-20,-20,-20)$ for the target structures (annotated in red and
        blue, resepectively), we demonstrate the
        effectivity of our adaptive multi-dimensional Boltzmann sampling
        procedure; moreover, for comparison, the distributions for uniform
        (U) and Boltzmann (B) samples---respectively associated with
        homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    {\sf \bfseries A}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/PKB00211_PKB00239_0_GC_content}
    ~~~~~{\sf \bfseries B}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/supp_3str_GC_content}
  \end{center}
  \caption{Distribution of the \GCb-content \GCb\% in the experiments of (A) Fig.~\ref{fig:energydist-pk} of the main text and (B) Suppl.~Fig.~\ref{fig:energydist}.}
  \label{fig:gc-content}
\end{figure}

\begin{revision}
\section{Effectivity of the seed sequence generation}

Figure~\ref{fig:seed-generation} shows relative numbers of finished instances and admissible sequences at each iteration of the seed generation procedure for the benchmark set \texttt{3str}. We obtain qualitatively equivalent results for all other benchmark sets (not shown). The top plot illustrates the  precomputation step of our strategy, where weights are calibrated for \GCb\% of 0.6. In this step, we produce sequence of the targeted \GCb\% to determine the linear transformation from the sampling energy model (stacking model) to the Turner model. The bottom plot shows the actual generation of the seed sequences after precomputation. The plot shows that most of the instances are finished after the first iterations, with only a few harder instances that require 2--3 times more iterations. Recall that the seed sequences were designed to have similar energy for all target structures.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.48\textwidth]{Figs/SeedGeneration/iterations_gc_3str.pdf}
  \includegraphics[width=0.48\textwidth]{Figs/SeedGeneration/iterations_en_3str.pdf}
  \caption{Relative numbers of finished instances and admissible sequences at each iteration of the seed generation procedure for the benchmark set \texttt{3str}.}
  \label{fig:seed-generation}
\end{figure}

@@

For validating the trained parameters, we generate a second
independent test set of random RNA sequences in the same
way. Fig.~\ref{fig:training-cor} shows correlation plots for the
trained parameters in the base pair and stacking models for predicting
the Turner energies in the test set with respective correlations of
$0.95$ an $0.94$.

\begin{figure}[h!]
  \centering
  A)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/basepaircor}
  B)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/stackingcor}
  \caption{Validation of the trained parameters for the A) base pair
    model B) stacking model for predicting energies of the
    independently sampled sequences in the test set. We show
    correlation plots against the Turner energies reported by the
    ViennaRNA package.}
  \label{fig:training-cor}
\end{figure}

@@

    \section{Exploiting constraint consistency to reduce the complexity}\label{sec:improvedComplexity}

@@

    %\section{Exploiting the bipartite nature of base pair dependency
    %  graphs to reduce the complexity}\label{sec:first-improvedComplexity}

@@

\section{Monotonicity of the partial derivatives within weight calibration}
\label{sec:weight_derivatives}
%
In weighted distributions, one witnesses a fairly predictable impact of the variation of any weight $\pi_\ell$ over the expected value $\mathbb{E}(E_\ell\mid \pmb{\pi})$, $\pmb{\pi}:= (\pi_1\cdots\pi_k)$, of the free energy $E_\ell$ of structure $\ell$.  Let us  first remind the probability of a sequence $S$ in the weighted distribution
\begin{align*}
  \mathcal{B}_{\pmb{\pi}}(S) &= \prod_{i=1}^{k} \pi_i^{-E_i(S)}, &
  \partfun{\pmb{\pi}}&=\sum_{S'}\mathcal{B}_{\pmb{\pi}}(S') &
    \text{and}& &
  \mathbb{P}(S\mid \pmb{\pi}) &= \frac{\mathcal{B}_{\pmb{\pi}}(S)}{\partfun{\pmb{\pi}}}.
  \end{align*}

@@

\section{Approximate counting and random generation}
In fact, not only is $\#{\sf BIS}$ a reference problem in counting complexity, but it is also a landmark problem with respect to the complexity of approximate counting problems. In this context, it is the representative for a class of $\#{\sf BIS}$-hard problems~\citep{Bulatov2013} that are easier to approximate than $\# {\sf SAT}$, yet are widely believed not to admit any Fully Polynomial-time Randomized Approximation Scheme. Recent results reveal a surprising dichotomy in the behavior of $\#{\sf BIS}$: it admits a Fully Polynomial-Time Approximation Scheme (FPTAS) for graphs of max degree $\le 5$~\citep{Weitz2006}, but is as hard to approximate as the general $\#{\sf BIS}$ problem on graphs of degree $\ge 6$~\citep{Cai2016}. In other words, there is a clear threshold, in term of the max degree, separating (relatively) easy instances from really hard ones.

@@

\section{Contribution of terms to the MultiDefect score}

@@

Figure~\ref{fig:contribtions-multidefect} breaks down the multi-defect
score of Eq.~$(\ref{eq:blueprintobjective})$, which is used to
evaluate the quality of multi-target designs. Recall that the total
score is composed of the term $\frac{1}{k} \sum_{\ell=1}^{k} (E(S,
R_\ell) - \EnsE(S))$, which evaluates the distances of the energies of
target structures to the minimum free energy (Distance to MFE), and
the term $\frac{1}{2\binom{k}{2}} \sum\limits_{1\leq\ell<j\leq
  k}|E(S,R_\ell) - E(S,R_j)|$, which evaluates the similarity of the
target structure energies of the target structures (Heterogeneity).

\label{appsec:multidefect}
\begin{figure}[h!]
\includegraphics[width=.48\textwidth]{Figs/statistics-Term1}
\includegraphics[width=.48\textwidth]{Figs/statistics-Term2}
\caption{Break-down of average MultiDefect values into their two components.}
\label{fig:contribtions-multidefect}
\end{figure}

\section{Detailed results of MultiDefect analysis}\label{sec:validity}
N/A values correspond to data that could not be obtained by the time of this submission for two main reasons: In the case of the initial sampling (left columns), they correspond to instances which, in conjunction with an expressive energy model, resulted in very high tree-width, leading to unreasonable memory requirements incompatible with our available computing resources; The case of missing values after optimization  (right columns), they indicate situations where the initial optimization took too long ($\approx$ 1 day) and was killed.
\input{table}

>>>>

\section{Immediate benefits of positive design for negative design}
\label{appsec:immediate-benefits-for-negative-design}
Already previous work on RNA Design, e.g. INFO-RNA and Incarnation, observed
that---put briefly---negative design benefits from positive design. Here, more
concretely, we simply \emph{demonstrate} this effect (for small examples), where
we consider \Def{negative design} as  
designing RNA sequences that fold with high probability $Pr(R|S)$ into the
given target structure $R$; and \Def{positive design}, as generating one or
many sequences that have a low energy for $R$.

@@

Note that low energy $E(S|R)$ is (non-linearly) correlated with high
probability $Pr(S|R)$, the probability to see $S$ in the Boltzmann distribution
of sequences folding into $R$. Recall that the probability $Pr(R|S)$ can be
calculated as $\exp(-E(S,R)/RT)/Z(S)$,
where $Z(S)=\sum_{R'} \exp(-E(S,R')/RT)$.
%
Consequently, Figure~\ref{appfig:negforpos} ultimately illustrates the dependencies
between $Pr(S|R)$ and $Pr(R|S)$ for uniform samples and low energy samples
(where those samples can be efficiently generated due to our Boltzmann sampling
strategy). For our examples, we observe (Fig.~\ref{appfig:negforpos}) that Boltzmann sampling with weight 500 (green), generates good negative designs 
with substantially higher probability than uniform sampling (blue).

@@

\begin{figure}[!h]
\textbf{A}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/stem_weight1-500_mean}%
\textbf{B}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight1-500_mean}%
\caption{Comparison of uniform sampling and Boltzmann sampling with
weight $\pi=500$. We show results for two different small structures: (A) single stem structure (B) double stem structure. For each respective structure $R$, we plot the energies $E(S,R)$ of
10000 sampled sequences against the probabilities $Pr(R|S)$ (in log scale). For illustrating the trend, we show moving averages over windows of $\pm 1$ kcal. Note that while only the energy $E(S,R)$ is directly controlled by Boltzmann sampling, this strongly impacts the common negative design objective $Pr(R|S)$.}
\label{appfig:negforpos}
\end{figure}

@@

\begin{figure}[h!]
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering\includegraphics[width=0.8\textwidth]{Figs/td-widths}
  \caption{Treewidths for multi-target RNA design instances of
    different complexity. Distributions of treewidths shown as boxplots for the base pair (left) and stacking model (right).}
  \label{appfig:td-widths}
\end{figure}

\begin{figure}[h!]
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering
  \includegraphics[width=0.8\textwidth]{Figs/td-times}
  \caption{Computation time of tree decompositions for
    multi-target RNA design instances of different complexity.
    Distributions of times (in ms/instance) shown as boxplots for the base pair (left) and stacking model (right).}
  \label{appfig:td-times}
\end{figure}

For each instance, we generated the dependency graphs in the base pair
model and in the stacking model. Then, we performed tree decomposition
(using strategy ``GreedyFillIn'' of LibTW~\citep{Dijk2006}) on each dependency
graph. The obtained treewidths are reported in
Fig.~\ref{appfig:td-widths}, while Fig.~\ref{appfig:td-times} shows the
corresponding run-times of the tree decomposition.

\begin{figure}[h!]
  %% the figure shows benchmark_inputs/RNAfold/3str/f3.100.0.inp
  \centering
  {\small \verbatiminput{Figs/f3_100_0.txt}}
  \includegraphicstop[width=0.55\textwidth]{Figs/td-example-basepair}%
  \includegraphicstop[width=0.30\textwidth]{Figs/td-example-stacking}
  \caption{Example instance from RF3 (top) with its tree decompositions in the base pair (left) and stacking model (right). The respective treewidths are 2 and 12.}
  \label{appfig:td-example}
\end{figure}

\revised{
  Finally, for further illustration, we show concrete tree decompositions for a selected instance in the benchmark set RF3 in Figure~\ref{appfig:td-example}. This figure is provided solely for illustrational purposes:
  First, it depicts the tree decomposition for a typical, realistic sampling task. Second, it exemplarily illustrates the typical impact of the energy model on the tree decomposition.
  Compared to the tree decomposition based on the stacking model, the tree decomposition for the base pair model is much simpler. This shows in many connected components, which correspond to independently solvable sub-problems, and fewer positions per bag (i.e. smaller tree width), which allows far more efficient evaluation per tree node. 
  Note that the complexity of the tree decomposition ultimately shows the complexity of the dependencies that need to be considered for simultaneous energy evaluation for all target structures. This complexity is substantially increased in the stacking model over the simpler base pair model.
}

@@

\section{Run times for Boltzmann sampling (at fixed weights)}
\label{appsec:run-times}

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times (Intel(R) Core(TM) i7-4770 CPU) for generating 10000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant pre-computation, it is informative to separately show construction and sampling times (Figure~\ref{appfig:run-times-redprint}). Moreover, \ourprog{} can be used with a simpler base pair model or a rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width and thus the runtime, as noticeable in Figure~\ref{appfig:run-times-redprint}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \ourprog{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences. }
  \label{appfig:run-times-redprint}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \RNAblueprint{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences.}
  \label{appfig:run-times-rbp}
\end{figure}

A fair comparison to sampling times by other tools can hardly be given, since the other tools produce samplings of very different nature. In particular, Modena does not make any guarantees on properties of the sampling. Non-surprisingly, this allows it to sample significantly faster (several orders of magnitudes).
Nevertheless, the comparison to run times of \RNAblueprint{} is still interesting, since this tool generates---similarly to our tool---\emph{controlled} samples. As we argue that the Boltzmann sampling is fundamentally more flexible than uniform sampling as generated by \RNAblueprint, \ourprog{} performs even particularly well, beating \RNAblueprint{} most of the time with only a few exceptions for particularly hard instances. The results for \RNAblueprint{} are shown in Figure~\ref{appfig:run-times-rbp}.

@@

\section{Illustrating multi-dimensional Boltzmann sampling for three pseudoknot-free target structures}
\label{appsec:illustrating-mdbs}

In analogy to Fig.~\ref{fig:energydist-pk} of the main text,
    Fig.~\ref{appfig:energydist} illustrates our multi-dimensional sampling
    strategy for three pseudoknot-free target structures (example instance
    from the \texttt{Modena} benchmark). The figure shows Turner energy
    distributions of single structures from sampling with different weight
    parameters, where weights are trained to target specific energies. For
    comparison, the figure shows distributions from uniform and Boltzmann
    sampling at respective homogenous weights $1$ and $e^\beta$.

    Fig.~\ref{appfig:gc-content} shows the distributions of the GC content of the samples, while targeting different combinations of energies in this experiment. For completeness, we also report corresponding results for the analogous experiment from the main text. We observe that in addition to the target energies, also the feature $\GCb\%$ is narrowly defined by the multi-dimensional Boltzmann sampling strategy. In comparison, for uniform sampling and Boltzmann sampling at high weight $e^\beta$, where the \GCb\% is not controlled, we obtain much broader distributions.

\begin{figure}[h!]
      \begin{center}
        %{\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
        %{\sf \bfseries B}
        \includegraphicstop[width=0.8\textwidth]{Figs/energy_distribution}
      \end{center}
      \caption{%
        Targeting specific energies using multi-dimensional Boltzmann
        sampling. We show the Turner energy distributions for
        three target structures R1, R2 and R3. By
        targeting the respective free energies $(-40,-40,-20)$ and
        $(-20,-20,-20)$ for the target structures (annotated in red and
        blue, resepectively), we demonstrate the
        effectivity of our adaptive multi-dimensional Boltzmann sampling
        procedure; moreover, for comparison, the distributions for uniform
        (U) and Boltzmann (B) samples---respectively associated with
        homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{appfig:energydist}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    {\sf \bfseries A}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/PKB00211_PKB00239_0_GC_content}
    ~~~~~{\sf \bfseries B}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/supp_3str_GC_content}
  \end{center}
  \caption{Distribution of the \GCb-content \GCb\% in the experiments of (A) Fig.~\ref{fig:energydist-pk} of the main text and (B) Suppl.~Fig.~\ref{appfig:energydist}.}
  \label{appfig:gc-content}
\end{figure}

\begin{revision}
\section{Effectivity of the seed sequence generation}
\label{appsec:effectivity-seed-generation}

Figure~\ref{appfig:seed-generation} shows relative numbers of finished instances and admissible sequences at each iteration of the seed generation procedure for the benchmark set \texttt{3str}. We obtain qualitatively equivalent results for all other benchmark sets (not shown). The first plot illustrates the  precomputation step of our strategy, where weights are calibrated for \GCb\% of 0.6. In this step, we produce sequence of the targeted \GCb\% to determine the linear transformation from the sampling energy model (stacking model) to the Turner model. The second plot shows the actual generation of the seed sequences after precomputation. The plot shows that most of the instances are finished after the first iterations, with only a few harder instances that require 2--3 times more iterations. Recall that the seed sequences were designed to have similar energy for all target structures.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.48\textwidth]{Figs/SeedGeneration/iterations_gc_3str.pdf}
  \includegraphics[width=0.48\textwidth]{Figs/SeedGeneration/iterations_en_3str.pdf}
  \caption{Relative numbers of finished instances and admissible sequences at each iteration of the seed generation procedure for the benchmark set \texttt{3str}. The left illustrates the procedure for targeting just the \GCb\% in the precomputation step; this step is used to learn instance-specific transformations of approximate energies to the Turner model from the generated sample. The right plots illustrates generating samples targeting \GCb\% and the Turner energies. }
  \label{appfig:seed-generation}
\end{figure}

@@

For validating the trained parameters, we generate a second
independent test set of random RNA sequences in the same
way. Fig.~\ref{appfig:training-cor} shows correlation plots for the
trained parameters in the base pair and stacking models for predicting
the Turner energies in the test set with respective correlations of
$0.95$ an $0.94$.

\begin{figure}[h!]
  \centering
  A)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/basepaircor}
  B)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/stackingcor}
  \caption{Validation of the trained parameters for the A) base pair
    model B) stacking model for predicting energies of the
    independently sampled sequences in the test set. We show
    correlation plots against the Turner energies reported by the
    ViennaRNA package.}
  \label{appfig:training-cor}
\end{figure}

@@

    \section{Exploiting constraint consistency to reduce the complexity}
    \label{appsec:improvedComplexity}

@@

    %section{Exploiting the bipartite nature of base pair dependency
    %  graphs to reduce the complexity}\label{appsec:first-improvedComplexity}

@@

\section{Monotonicity of the partial derivatives within weight calibration}
\label{appsec:weight_derivatives}

%
In weighted distributions, one witnesses a fairly predictable impact of the variation of any weight $\pi_\ell$ over the expected value $\mathbb{E}(E_\ell\mid \pmb{\pi})$, $\pmb{\pi}:= (\pi_1\cdots\pi_k)$, of the free energy $E_\ell$ of structure $\ell$.  Let us  first remind the probability of a sequence $S$ in the weighted distribution
\begin{align*}
  \mathcal{B}_{\pmb{\pi}}(S) &= \prod_{i=1}^{k} \pi_i^{-E_i(S)}, &
  \partfun{\pmb{\pi}}&=\sum_{S'}\mathcal{B}_{\pmb{\pi}}(S') &
    \text{and}& &
  \mathbb{P}(S\mid \pmb{\pi}) &= \frac{\mathcal{B}_{\pmb{\pi}}(S)}{\partfun{\pmb{\pi}}}.
  \end{align*}

@@

\section{Approximate counting and random generation}
\label{appsec:approximate-counting}

In fact, not only is $\#{\sf BIS}$ a reference problem in counting complexity, but it is also a landmark problem with respect to the complexity of approximate counting problems. In this context, it is the representative for a class of $\#{\sf BIS}$-hard problems~\citep{Bulatov2013} that are easier to approximate than $\# {\sf SAT}$, yet are widely believed not to admit any Fully Polynomial-time Randomized Approximation Scheme. Recent results reveal a surprising dichotomy in the behavior of $\#{\sf BIS}$: it admits a Fully Polynomial-Time Approximation Scheme (FPTAS) for graphs of max degree $\le 5$~\citep{Weitz2006}, but is as hard to approximate as the general $\#{\sf BIS}$ problem on graphs of degree $\ge 6$~\citep{Cai2016}. In other words, there is a clear threshold, in term of the max degree, separating (relatively) easy instances from really hard ones.

@@

\section{Contribution of terms to the MultiDefect score}
\label{appsec:contributions-multidefect}

@@

Figure~\ref{appfig:contributions-multidefect} breaks down the multi-defect
score of Eq.~$(\ref{eq:blueprintobjective})$, which is used to
evaluate the quality of multi-target designs. Recall that the total
score is composed of the term $\frac{1}{k} \sum_{\ell=1}^{k} (E(S,
R_\ell) - \EnsE(S))$, which evaluates the distances of the energies of
target structures to the minimum free energy (Distance to MFE), and
the term $\frac{1}{2\binom{k}{2}} \sum\limits_{1\leq\ell<j\leq
  k}|E(S,R_\ell) - E(S,R_j)|$, which evaluates the similarity of the
target structure energies of the target structures (Heterogeneity).

\begin{figure}[h!]
\includegraphics[width=.48\textwidth]{Figs/statistics-Term1}
\includegraphics[width=.48\textwidth]{Figs/statistics-Term2}
\caption{Break-down of average MultiDefect values into their two components.}
\label{appfig:contributions-multidefect}
\end{figure}

\section{Detailed results of the negative design benchmark}
\label{appsec:detailed-benchmark-results}

N/A values correspond to data that could not be obtained by the time of this submission for two main reasons: In the case of the initial sampling (left columns), they correspond to instances which, in conjunction with an expressive energy model, resulted in very high tree-width, leading to unreasonable memory requirements incompatible with our available computing resources; The case of missing values after optimization  (right columns), they indicate situations where the initial optimization took too long ($\approx$ 1 day) and was killed.

\input{table}
